---
title: "BIO 777: Data Wrangling and Analysis Using R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Outline for today

0. Caveats: I'm an R newbie too! Mostly. Sorta?
1. Go over syllabus
2. Install R and RStudio
3. What is R?
4. What are Markdown and R Markdown?

# Syllabus

- The syllabus will be located here: <https://github.com/vancleve/BIO770-DWVR>. 
- This is also where R and R Markdown documents will be located. The only files that will be kept on the [Canvas site](https://uk.instructure.com/courses/1843807) are the PDF copies of the reference books and the completed labs and figures that you submit.
- I'll update the syllabus and schedule of topics as necessary.

# Install R and RStudio

## Downloading

If you haven't already, download R and R studio

- To download R, go here: <https://mirrors.nics.utk.edu/cran/>
- To download RStudio, go here <https://www.rstudio.com/products/rstudio/download/preview/>
    - Note: this is the "preview version", which is preferable. The "preview version" includes something called an "R Notebook", which is basically a more interactive way to use R Markdown. I'll be using this during the class.

## Installing

If you need directions, especially for how update your packages if you already have R, go here:
<http://stat545.com/block000_r-rstudio-install.html>

Otherwise, proceed with your installing normal if you haven't already.

## Running RStudio

### What is RStudio?

RStudio is an Integrated Development Environment (IDE) for the R progamming language. While you can write R code in any text editor you like and then run that code with the R interpreter, there are many things that an IDE can do that help you be more efficient when programming.

1. Synatx highlighting (e.g., see the markdow!)
2. File/project organization (see "Files" pane)
3. Examining variables that you've set (see the "Environment" pane)
4. Easily execute code and examine its text output (see "Console") or graphical output (see "Viewer")
5. View help files (see "Help")

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
